---
title: "DS4420 Developer Distrust Project - Bayes"
author: "Gavin Bond"
date: "'r Sys.Date()"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(mc.cores = min(4, parallel::detectCores()))
set.seed(42)
```

Necessary imports:
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(brms)
```

Initial data import (2 survey years combined). Column types 
specified due to parsing issue on raw import.
```{r}
df <- readr::read_csv(
  "data/combined_survey.csv",
  col_types = readr::cols(.default = readr::col_character()),
  show_col_types = FALSE
)
```

Ensuring parsing issues were fixed
```{r}
problems(df)
```

Data inspection 1 (dimensions, top 5 rows, cols)
```{r}
dim(df)
head(df)
names(df)
```

Data inspection 2 (summary, checking for NAs)
```{r}
summary(df)

table(df$AIAcc, useNA = "ifany")
table(df$AISelect, useNA = "ifany")
table(df$year, useNA = "ifany")

colSums(is.na(df))
```
Helper function for preprocessing years of experience
```{r}
# Helper
clean_experience <- function(x) {
  x <- trimws(x)
  
  dplyr::case_when(
    is.na(x) ~ NA_real_,
    x == "" ~ NA_real_,
    x == "Less than 1 year" ~ 0.5,
    x == "More than 50 years" ~ 51,
    TRUE ~ suppressWarnings(as.numeric(x))
  )
}

# Implementing helper
clean_factor <- function(x) {
  x <- trimws(x)
  x[is.na(x) | x == ""] <- "Missing"
  as.factor(x)
}
```

One more inspection block for messy numeric fields
```{r}
sort(unique(df$YearsCode))[1:10]
tail(sort(unique(df$YearsCode)), 10)

sort(unique(df$WorkExp))[1:10]
tail(sort(unique(df$WorkExp)), 10)
```
Initial Bayes model processing:
- Restricting model to AI users only
- Predicting distrust vs. trust
- Using simple feature set for first attempt
```{r}
df_model <- df %>%
  dplyr::filter(AISelect == "Yes") %>%
  dplyr::transmute(
    y = ifelse(AIAcc %in% c("Somewhat distrust", "Highly distrust"), 1, 0),
    YearsCode = clean_experience(YearsCode),
    Age = clean_factor(Age),
    EdLevel = clean_factor(EdLevel),
    OrgSize = clean_factor(OrgSize),
    AISent = clean_factor(AISent),
    AIThreat = clean_factor(AIThreat),
    RemoteWork = clean_factor(RemoteWork),
    AIComplex = clean_factor(AIComplex),
    year = clean_factor(year)
  )

dim(df_model)
head(df_model)
table(df_model$y)
```

Final preprocessing for Bayes model
```{r}
summary(df_model$YearsCode)
sum(is.na(df_model$YearsCode))

df_model <- df_model %>%
  dplyr::filter(!is.na(YearsCode))

df_model$YearsCode_z <- as.numeric(scale(df_model$YearsCode))

summary(df_model$YearsCode_z)
dim(df_model)
table(df_model$y)
```

Final modeling df
```{r}
df_model <- df_model %>%
  dplyr::select(
    y, YearsCode_z, Age, EdLevel, OrgSize,
    AISent, AIThreat, RemoteWork, AIComplex, year
  )

str(df_model)
```

Initial bayes log-regression formula
```{r}
bayes_formula <- bf(
  y ~ YearsCode_z + Age + EdLevel + OrgSize +
    AISent + AIThreat + RemoteWork + AIComplex + year
)

brms::default_prior(
  bayes_formula,
  data = df_model,
  family = bernoulli("logit")
)
```