---
title: "DS4420 Developer Distrust Project - Bayes"
author: "Gavin Bond"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(mc.cores = min(4, parallel::detectCores()))
set.seed(42)
```

Necessary imports:
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(brms)
```

Initial data import (2 survey years combined). Column types 
specified due to parsing issue on raw import.
```{r}
df <- readr::read_csv(
  "data/combined_survey.csv",
  col_types = readr::cols(.default = readr::col_character()),
  show_col_types = FALSE
)
```

Ensuring parsing issues were fixed
```{r}
problems(df)
```

Data inspection 1 (dimensions, top 5 rows, cols)
```{r}
dim(df)
head(df)
names(df)
```

Data inspection 2 (summary, checking for NAs)
```{r}
summary(df)

table(df$AIAcc, useNA = "ifany")
table(df$AISelect, useNA = "ifany")
table(df$year, useNA = "ifany")

colSums(is.na(df))
```
From this, we can see there are many missing values; in terms of how it will
effect our modeling choices, we are intentionally removing DevType, WorkExp, 
and ICorPM for our initial fit due to high levels of missing values, high
level count, and feature overlap. Will reevaluate this decision after full 
first model implementation.

Helper function for preprocessing years of experience
```{r}
# Helper for experience
clean_exp <- function(x) {
  x <- trimws(x)
  out <- dplyr::case_when(
    is.na(x) ~ NA_real_,
    x == "" ~ NA_real_,
    x == "Less than 1 year" ~ 0.5,
    x == "More than 50 years" ~ 51,
    TRUE ~ suppressWarnings(as.numeric(x))
  )
  pmin(out, 51)
}

# Helper for other missing factor values
clean_factor <- function(x) {
  x <- trimws(x)
  x[is.na(x) | x == ""] <- "Missing"
  as.factor(x)
}
```

Final inspection for numeric fields
```{r}
sort(unique(df$YearsCode))[1:10]
tail(sort(unique(df$YearsCode)), 10)

sort(unique(df$WorkExp))[1:10]
tail(sort(unique(df$WorkExp)), 10)
```
Initial model processing:
- Restricting model to AI users only
- Predicting distrust vs. non-distrust (non-distrust includes "trust" and "neutral")
- Using simple feature set for first attempt
```{r}
df_model <- df %>%
  dplyr::filter(AISelect == "Yes") %>%
  dplyr::transmute(
    y = ifelse(AIAcc %in% c("Somewhat distrust", "Highly distrust"), 1, 0),
    YearsCode = clean_exp(YearsCode),
    Age = clean_factor(Age),
    EdLevel = clean_factor(EdLevel),
    OrgSize = clean_factor(OrgSize),
    AISent = clean_factor(AISent),
    AIThreat = clean_factor(AIThreat),
    RemoteWork = clean_factor(RemoteWork),
    AIComplex = clean_factor(AIComplex),
    year = clean_factor(year)
  )

dim(df_model)
head(df_model)
table(df_model$y)
```

Final preprocessing for model
```{r}
summary(df_model$YearsCode)
sum(is.na(df_model$YearsCode))

df_model <- df_model %>%
  dplyr::filter(!is.na(YearsCode))

df_model$YearsCode_z <- as.numeric(scale(df_model$YearsCode))

summary(df_model$YearsCode_z)
dim(df_model)
table(df_model$y)
```

Final modeling df
```{r}
df_model <- df_model %>%
  dplyr::select(
    y, YearsCode_z, Age, EdLevel, OrgSize,
    AISent, AIThreat, RemoteWork, AIComplex, year
  )

str(df_model)
```

Initial bayes log-regression formula
```{r}
init_bayes_formula <- bf(
  y ~ YearsCode_z + AISent + AIThreat + AIComplex + year
)

# Default prior
init_def_prior <- default_prior(
  init_bayes_formula,
  data = df_model,
  family = bernoulli("logit")
)

init_def_prior

# Manual prior
init_manual_prior <- c(
  brms::set_prior("student_t(4, 0, 10)", class = "Intercept"),
  brms::set_prior("normal(0, 2)", class = "b")
)
```

Fitting initial bayes log-regression model 
```{r}
init_brm_logit <- brms::brm(
  formula = init_bayes_formula,
  family = bernoulli("logit"),
  data = df_model,
  chains = 4,
  iter = 1000,
  warmup = 180,
  prior = init_manual_prior,
  seed = 4420
)
```

Checking initial model results & diagnostics
```{r}
summary(init_brm_logit)
```
Checking initial model diagnostics plot
```{r}
#plot(init_brm_logit)
```

Posterior predicted probabilities for initial model
```{r}
# Computing predictions
init_pred_probs <- fitted(init_brm_logit)[, "Estimate"]

# Viewing prediction sample and summary
head(init_pred_probs)
summary(init_pred_probs)
```
Improved summary output and plots
```{r}
fixef(init_brm_logit)

plot(conditional_effects(init_brm_logit, "YearsCode_z"))
plot(conditional_effects(init_brm_logit, "AISent"))
plot(conditional_effects(init_brm_logit, "AIThreat"))
plot(conditional_effects(init_brm_logit, "AIComplex"))
```
Predicted probabilities by observed class
```{r}
prob_df <- data.frame(
  prob = init_pred_probs,
  y = as.factor(df_model$y)
)

# New plot, posterior mean prob. vs actual count
ggplot(prob_df, aes(x = prob, fill = y)) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30) +
  labs(
    title = "Predicted Probability of Distrust by Observed Class",
    x = "Posterior Mean Probability of Distrust",
    y = "Count",
    fill = "Observed y"
  )
```

Boxplot of predicted probabilities by class
```{r}
ggplot(prob_df, aes(x = y, y = prob)) +
  geom_boxplot() +
  labs(
    title = "Predicted Probability of Distrust by Observed Class",
    x = "Observed Distrust",
    y = "Posterior Mean Probability of Distrust"
  )
```

Building potential poster visual 1
```{r}
coef_df <- as.data.frame(fixef(init_brm_logit))
coef_df$term <- rownames(coef_df)

coef_plot_df <- coef_df %>%
  dplyr::filter(term != "Intercept") %>%
  dplyr::filter(!grepl("Missing", term)) %>%
  dplyr::mutate(
    term = dplyr::case_when(
      term == "YearsCode_z" ~ "Years of Coding (z)",
      term == "AISentIndifferent" ~ "AI Sentiment: Indifferent",
      term == "AISentUnfavorable" ~ "AI Sentiment: Unfavorable",
      term == "AISentUnsure" ~ "AI Sentiment: Unsure",
      term == "AISentVeryfavorable" ~ "AI Sentiment: Very favorable",
      term == "AISentVeryunfavorable" ~ "AI Sentiment: Very unfavorable",
      term == "AIThreatNo" ~ "AI Threat: No",
      term == "AIThreatYes" ~ "AI Threat: Yes",
      term == "AIComplexGoodbutnotgreatathandlingcomplextasks" ~ "AI Complex: Good, not great",
      term == "AIComplexIdontuseAItoolsforcomplextasksDIdontknow" ~ "AI Complex: Don't use / don't know",
      term == "AIComplexNeithergoodorbadathandlingcomplextasks" ~ "AI Complex: Neither good nor bad",
      term == "AIComplexVerypoorathandlingcomplextasks" ~ "AI Complex: Very poor",
      term == "AIComplexVerywellathandlingcomplextasks" ~ "AI Complex: Very well",
      term == "year2025" ~ "Year: 2025",
      TRUE ~ term
    )
  ) %>%
  dplyr::arrange(Estimate)

coef_plot <- ggplot(coef_plot_df, aes(x = Estimate, y = reorder(term, Estimate))) +
  geom_point() +
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    title = "Bayesian Logistic Regression Coefficients",
    x = "Posterior Mean Log-Odds Effect",
    y = NULL
  )

coef_plot
ggsave("figures/bayes_coef_forest.png", coef_plot, width = 9, height = 6, dpi = 300)
```

Building potential poster visual 2:
```{r}
# conditional effects
ce_aisent <- conditional_effects(init_brm_logit, "AISent")
# Df for reordering of factors
aisent_df <- ce_aisent$AISent

# Reordering factor levels for better layour
aisent_df$AISent <- factor(
  aisent_df$AISent,
  levels = c(
    "Very favorable",
    "Favorable",
    "Indifferent",
    "Unsure",
    "Unfavorable",
    "Very unfavorable",
    "Missing"
  )
)

p_aisent <- ggplot(aisent_df, aes(x = AISent, y = estimate__)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower__, ymax = upper__), width = 0.2) +
  labs(
    title = "Predicted Distrust by AI Sentiment",
    x = "AI Sentiment",
    y = "Predicted Probability of Distrust"
  )
p_aisent 

ggsave("figures/bayes_conditional_aisent.png", p_aisent, width = 8, height = 5, dpi = 300)
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```